---
title: "Conserve VA Refinement Method"
author: "Emma Jones"
date: "1/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# run in R 4.0.3

library(tidyverse)
library(sf)
library(readxl)
```

## Background

This script takes the initial DEQ results from the ConserveVA Methodolgy.Rmd script and refines results to a scale more desired by DCR. The data inputs are:
* Sites selected from original DEQ Conserve VA methodology
* Watersheds selected from original DEQ Conserve VA methodology
* HUC12 watershed layer
* 1:24k NHD

### Method

To limit area inclusions in Conserve VA more in line with DCR's methods, DEQ will:
1. Intersect the delineated watershed with the HUC12 layer.
    + Where the delineated watershed is larger than the HUC12 polygon that includes the station, only the HUC12 polygon the station falls into will be used for further analyses.
2. The resulting polygon will serve as the boundary to clip 1:24k NHD.
3. The 1:24k stream segments will be buffered to a set 200ft buffer on either side of the segment and dissolved into a single polygon.

### Output

DEQ will provide DCR with:
* Polygon layer of 1:24k stream segments within delineated watershed or HUC12 high quality station falls into (whichever is smaller) buffered to a set 200ft width
* Polyline layer of 1:24k stream segments within delineated watershed or HUC12 high quality station falls into (whichever is smaller) in case DCR wants to run segments through their stream buffering model.


### Methodology

Bring in original final stations and watersheds.

```{r original final datasets}
sites <- st_read('Results/forDCR/sites.shp') %>% arrange(UID)
watersheds <- st_read('Results/forDCR/watersheds.shp') %>% arrange(UID)
```



Bring in HUC12 watershed layer and 1:24k NHD.

```{r HUC12 and NHD}
huc12 <- st_read('C:/HardDriveBackup/GIS/GIS_BaseLayer_Datasets.gdb', layer = 'HUC12_50_mile_buffer') %>%
  mutate(geometry = SHAPE) %>% st_drop_geometry() %>%
  st_as_sf()
nhd <- st_read('C:/HardDriveBackup/GIS/NHD_H_51_GDB.gdb', layer = 'NHDFlowline') %>%
  st_transform(st_crs(huc12))
  #st_read('C:/HardDriveBackup/GIS/GIS_BaseLayer_Datasets.gdb', layer = 'va_2014_aus_riverine') # 1:24k resolution but not official layer
```

Intersect sites with HUC12 layer.

```{r watershed huc12 intersection}
sitesHUC12 <- st_intersection(sites, huc12) %>%#st_buffer(huc12, 0))#st_buffer(watersheds[i,],0)
  arrange(UID)
```

Sidebar, first try to get the final area desired bc R is having issues with random watersheds after 6 tries. Just going to clip nhd in gis to R determined area and buffer in GIS.

```{r}
finalArea <- watersheds[1,] %>%
  mutate(UID = 'TEST')

for(i in 1:nrow(sitesHUC12)){
  print(i)
  # find area 
  watershedArea <- filter(watersheds, UID %in% sitesHUC12[i,]$UID)
  huc12Area <- filter(huc12, HUC12 %in% sitesHUC12[i,]$HUC_12)
  # choose smaller area
  if(st_area(watershedArea) > st_area(huc12Area)){
    chosenArea <- st_intersection(st_buffer(huc12Area, 0), st_buffer(watershedArea, 0)) %>% # clip HUC12 to watershed to remove part of HUC12 downstream of site
      mutate(UID = sitesHUC12[i,]$UID) %>%
      dplyr::select(UID)
    } else {
      chosenArea <- dplyr::select(watershedArea, UID)
    }
    finalArea <- bind_rows(finalArea, chosenArea)
}

finalArea <- filter(finalArea, UID != 'TEST') %>%
  dplyr::select(UID)


# now intersect nhd
flowlines <- st_intersection(st_buffer(finalArea, 0), st_zm(nhd))

# now buffer flowlines      
flowlinesBuffer <- st_transform(flowlines, crs= "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% # transform crs to get units
  st_buffer(dist = 60.96) # buffer 200ft (convert to m to match units)

# dissolve boundaries between polygons
flowlinesBufferSmooth <- flowlinesBuffer %>% 
  group_by(UID) %>%
  summarise() %>%
  st_transform(4326) # transform back to original crs
      
# clip buffered area to within original bounding polygon (to drop bit of buffer outside original poly downstream of site)
flowlinesBufferSmoothFin <- st_intersection(flowlinesBufferSmooth, st_buffer(finalArea, 0)) %>%
      dplyr::select(UID)

st_write(finalArea, 'Results/forDCR/finalArea.shp')
st_write(flowlines, 'Results/forDCR/flowlines.shp')
st_write(flowlinesBuffer, 'Results/forDCR/flowlinesBuffer.shp')
st_write(flowlinesBufferSmooth, 'Results/forDCR/flowlinesBufferSmooth.shp')
st_write(flowlinesBufferSmoothFin, 'Results/forDCR/flowlinesBufferSmoothFin.shp')
st_write(flowlinesBufferSmoothFin, 'Results/forDCR/riparian200ftBuffer.shp')

# remove duplicates for smaller area
flowlinesTotalArea <- mutate(flowlinesBufferSmooth, total = 'all') %>%
  group_by(total) %>%
  summarise()

sum(st_area(flowlinesTotalArea))/1000000 # m2 to km2

```







































Previous methods attempted (do not use):



First, remove all nested subwatersheds identified in previous analyses to remove duplicated data.

```{r nested watersheds}
#nestedSubwatersheds <- read_excel('Results/forDCR/Results.xlsx', sheet = 'nested_Subwatersheds') %>%
#  pivot_longer(-`Downstream Watershed`, names_to = 'Subwatershed', values_to = 'Station') %>%
#  filter(!is.na(Station))

#sites <- filter(sites, ! UID %in% nestedSubwatersheds$Station) 
#watersheds <- filter(watersheds, ! UID %in% nestedSubwatersheds$Station) 
```



Determine if intersected HUC12 or delineated watershed is larger.

- first test on small dataset to make sure methods work

```{r testing methods}
huc12Area <- filter(huc12, HUC12 %in% sitesHUC12[1,]$HUC_12)

st_intersection(huc12Area, st_zm(nhd))
```
- WKB type 11 issue, have a look at data structure

```{r troubleshoot issue}
as.data.frame(table(st_geometry_type(nhd)))
```

There is a MULTICURVE thrown in with MULTILINESTRING, cast the data to fix the geometry issue so geoprocessing can occur.

```{r cast data}
nhd2 <- st_cast(nhd, "MULTILINESTRING")
```

Now that everything is cast to the same data structure, try intersection again.

```{r test again}
st_intersection(huc12Area, nhd2)
```

Worked. Now remove extra data.

```{r testing cleanup}
rm(nhd); rm(huc12Area)
```



```{r}
finalArea <- watersheds[1,] %>%
  mutate(UID = 'TEST')

for(i in 1:nrow(sitesHUC12)){
  print(i)
  # find area 
  watershedArea <- filter(watersheds, UID %in% sitesHUC12[i,]$UID)
  huc12Area <- filter(huc12, HUC12 %in% sitesHUC12[i,]$HUC_12)
  # choose smaller area
  if(st_area(watershedArea) > st_area(huc12Area)){
    flowlines <- st_intersection(st_buffer(huc12Area, 0), st_zm(nhd)) %>%#nhd2) %>%
      # transform crs to get units
      st_transform(crs= "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m
                   +no_defs") %>%
      # buffer 200ft (convert to m to match units)
      st_buffer(dist = 60.96) %>%
      # add unifying feature for dissolve
      mutate(UID = sitesHUC12[i,]$UID) %>%
      # dissolve boundaries between polygons
      group_by(UID) %>%
      summarise() %>%
      # transform back to original crs
      st_transform(4326) %>%
      # clip buffered area to within original bounding polygon (to drop bit of buffer outside original poly downstream of site)
      st_intersection(st_buffer(huc12Area, 0)) %>%
      dplyr::select(UID)
    #flowlines <- st_cast(flowlines, 'MULTIPOLYGON')
    } else {
      flowlines <- st_intersection(st_buffer(watershedArea, 0), st_zm(nhd)) %>%#nhd2) %>%
        # transform crs to get units
        st_transform(crs= "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m
                     +no_defs") %>%
        # buffer 200ft (convert to m to match units)
        st_buffer(dist = 60.96) %>%
        # dissolve boundaries between polygons
        group_by(UID) %>%
        summarise() %>%
        # transform back to original crs
        st_transform(4326) %>%
        # clip buffered area to within original bounding polygon (to drop bit of buffer outside original poly downstream of site)
        st_intersection(st_buffer(watershedArea, 0)) %>%
        dplyr::select(UID)
      #flowlines <- st_cast(flowlines, 'MULTIPOLYGON')
    }
  #if(nrow(flowlines) > 0){
    finalArea <- bind_rows(finalArea, flowlines)
  #}
}

finalArea <- filter(finalArea, UID != 'TEST') %>%
  dplyr::select(UID)

#st_write(finalArea, 'finalArea5.shp')
```
```{r}
filter(finalArea, geometry == list())
```



Final area

```{r}
sum(st_area(finalArea))/1000000 # m2 to km2
```



z <- st_transform(flowlines, crs= "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>%
  st_buffer(dist = 60.96) %>%
  st_union()
plot(z)
```


Database:	C:\HardDriveBackup\GIS\GIS_BaseLayer_Datasets.gdb
Feature Dataset:	Hydrology
Feature Class:	HUC12_50_mile_buffer
